---
title: "R4DS Book Club - Asia"
subtitle: "Chapter 18 - Pipes"
author: "Adithi R. Upadhya, ILK Labs"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader
---


## Why is clean codes important?

> Even if you're not working with other people, you'll definitely be working with future-you! Writing clear code is important so that others (like future-you) can understand why you tackled an analysis in the way you did. That means getting better at programming also involves getting better at communicating. Over time, you want your code to become not just easier to write, but easier for others to read.
 
 
---
## Pipes

> The pipe, `%>%`, comes from the **magrittr** package by Stefan Milton Bache. Packages in the tidyverse load `%>%` for you automatically, so you don’t usually load magrittr explicitly. 


```{r warning = FALSE, message = FALSE}
library(magrittr)
library(pryr)
```
 
---

## Piping alternatives

> Little bunny Foo Foo

> Went hopping through the forest

> Scooping up the field mice

> And bopping them on the head

```{r eval = FALSE}
foo_foo_1 <- hop(foo_foo, through = forest)
foo_foo_2 <- scoop(foo_foo_1, up = field_mice)
foo_foo_3 <- bop(foo_foo_2, on = head)

```


---
 ## Downsides
 
 - Name each intermediate element.
 - Adding numeric suffixes are difficult to follow.
 - The code is cluttered with unimportant names.
 - You have to carefully increment the suffix on each line.
 - Memory however is not an issue, R is smart enough to share columns across data frames. 
 
```{r}
diamonds <- ggplot2::diamonds
diamonds2 <- diamonds %>% 
  dplyr::mutate(price_per_carat = price / carat)

pryr::object_size(diamonds)
pryr::object_size(diamonds2)
pryr::object_size(diamonds, diamonds2)

```
 
---
## Overwrite the original
 
```{r eval = FALSE}
foo_foo <- hop(foo_foo, through = forest)
foo_foo <- scoop(foo_foo, up = field_mice)
foo_foo <- bop(foo_foo, on = head)
```

- Debugging is painful: if you make a mistake you’ll need to re-run the complete pipeline from the beginning.
- The repetition of the object being transformed (we’ve written foo_foo six times!) obscures what’s changing on each line.

---
## Function composition

- Just string the function calls together.
- But this approach is difficult for a human to read.

```{r eval = FALSE}
bop(
  scoop(
    hop(foo_foo, through = forest),
    up = field_mice
  ), 
  on = head
)
```

 
---
## Use the pipe
 
```{r eval = FALSE}
foo_foo %>%
  hop(through = forest) %>%
  scoop(up = field_mice) %>%
  bop(on = head)
```


- How does pipe work?

```{r eval = FALSE}
my_pipe <- function(.) {
  . <- hop(., through = forest)
  . <- scoop(., up = field_mice)
  bop(., on = head)
}
my_pipe(foo_foo)
```

---
## Pipe cannot be used in two cases -

- Functions that use the current environment. For example, `assign()` will create a new variable with the given name in the current environment. The use of `assign()` with the pipe does not work here because it assigns it to a temporary environment used by `%>%`

```{r}
assign("x", 10)
x
"x" %>% assign(100)
x
env <- environment()
"x" %>% assign(100, envir = env)
x
```
---
 
- Functions that use lazy evaluation. In R, function arguments are only computed when the function uses them, not prior to calling the function. The pipe computes each element in turn, so you can’t rely on this behaviour.



```{r}
tryCatch(stop("!"), error = function(e) "An error")
stop("!") %>% 
  tryCatch(error = function(e) "An error")
```

---

## When not to use the pipe

> Your pipes are longer than (say) ten steps. In that case, create intermediate objects with meaningful names. That will make debugging easier, because you can more easily check the intermediate results, and it makes it easier to understand your code, because the variable names can help communicate intent.

> You have multiple inputs or outputs. If there isn’t one primary object being transformed, but two or more objects being combined together, don’t use the pipe.

> You are starting to think about a directed graph with a complex dependency structure. Pipes are fundamentally linear and expressing complex relationships with them will typically yield confusing code.

---

## Other tools from magrittr

> When working with more complex pipes, it’s sometimes useful to call a function for its side-effects. Maybe you want to print out the current object, or plot it, or save it to disk. Many times, such functions don’t return anything, effectively terminating the pipe.

> To work around this problem, you can use the "tee" pipe. `%T>%` works like %>% except that it returns the left-hand side instead of the right-hand side. It’s called "tee" because it’s like a literal T-shaped pipe.

```{r}
rnorm(100) %>%
  matrix(ncol = 2) %>%
  plot() %>%
  str()

rnorm(100) %>%
  matrix(ncol = 2) %T>%
  plot() %>%
  str()
```

---

> If you’re working with functions that don’t have a data frame based API
(i.e. you pass them individual vectors, not a data frame and expressions to be evaluated in the context of that data frame), you might find `%$%` useful. It "explodes" out the variables in a data frame so that you can refer to them explicitly. This is useful when working with many functions in base R:

```{r}
mtcars %$%
  cor(disp, mpg)
```

---

> For assignment magrittr provides the %<>% operator which allows you to replace code like:

```{r}
mtcars <- mtcars %>% 
  transform(cyl = cyl * 2)

mtcars %<>% transform(cyl = cyl * 2)
```

