---
title: "Workflow Basics and Data Transformation"
author: "Adithi R Upadhya"
date: "11/3/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coding Basics

- Use RStudio’s keyboard shortcut: `Alt` + `-` (the minus sign). Notice that RStudio automagically surrounds `<-` with spaces, which is a good code formatting practice. 
- Code is miserable to read on a good day, so giveyoureyesabreak and use spaces.
- Object names must start with a letter, and can only contain letters, numbers, `_` and `.`. You want your object names to be descriptive, so you’ll need a convention for multiple words. We recommend snake_case where you separate lowercase words with `_`.

- `i_use_snake_case`

- `otherPeopleUseCamelCase`

- `some.people.use.periods`

- `And_aFew.People_RENOUNCEconvention`

- Typos matter. Case matters.
- If you make an assignment, you don't get to see the value. You're then tempted to immediately double-check the result:

```{r}
y <- seq(1, 10, length.out = 5)
y
```

- This common action can be shortened by surrounding the assignment with parentheses, which causes assignment and "print to screen" to happen.

```{r}
(y <- seq(1, 10, length.out = 5))
```

- Press Alt + Shift + K. What happens? How can you get to the same place using the menus?
- This gives a menu with keyboard shortcuts. This can be found in the menu under `Tools -> Keyboard Shortcuts Help`.

```{r message = FALSE, warning =  FALSE}
library(nycflights13)
library(tidyverse)
```





